%main for analyse

\part{System Analysis}
\label{system_analysis}

\chapter{Description of Water Supply Systems}
\label{description_of_water_supply_systems}

 \input{report/analysis/sys_overview}

\input{report/analysis/setup_considerations}

\input{report/requirements/requirements_main}

\chapter{Network simplification}
\label{network_simplification}

\emph{This chapter gives a general introduction about the need for model simplification in WSSs. Different approaches and methods are discussed in a state of art manner, including the development and methodologies employed in this field. Lastly, an attempt is made to simplify the Randers WSS, using one of the techniques which are being introduced in this chapter. }

\section{Purpose of the model reduction}
\label{purpose_of_the_model_reduction}

As it was explained in \chapref{description_of_water_supply_systems}, the typical components of WSSs are reservoirs, pipes, pumps and valves. Each of these interconnected elements are dependant on their neighbours, thus the behaviour of the entire WSS depends on each of its elements. For simulation purposes, it is required that the model of the real-life network consists of thousands of elements in order to accurately replicate hydraulic behaviour and the topographical layout of the system. Such models are appropriate for simulation purposes, however, online optimisation tasks are much more computationally demanding, therefore simplified models are required. There are different approaches for model reduction, however the outcome of most of these methods is a hydraulic model with a smaller number of components than the original one. 

The use of Geographic Information Systems(GIS) in the water industry resulted in an increasing amount of information about actual network topology and service that can be utilized in a model \cite{johnson2016geographic}. As a consequence this, normally the simulation model of WSSs include exactly the same amount of components as in real-life, and therefore a large-scale are considered. 

The efficiency and accuracy of large-scale system reduction is highly dependant on the complexity of models and the selected method. In this field, several researches has been focusing on the verification of different methods and therefore provided case studies and results. 

\section{State of the art model reduction analysis}
\label{state_of_the_art_model_reduction_analysis}

In the frame of this project, the WSS in Randers is considered as a large-scale system and attempted to be reduced. However, before discussing the reduction on the actual model network, first the different methods for model reduction are introduced in a state of the art manner. It is worth noting that throughout the report, the terms reduction and simplification are used alternately as a description of the process of achieving a hydraulic model with a smaller number of components than the original. The different reduction approaches, discussed in this report are: 

\begin{itemize}
\item Skeletonization, which is researched in \cite{walski2003advanced}, \cite{MARTINEZSOLANO2017286} and \cite{battlepaper}.
\item Parameter fitting (cite)
\item Graph decomposition (cite)
\item Variable elimination (cite)
\end{itemize}

\subsection{Skeletonization}
\label{skeletonization}

Skeletonization can be considered as a reduction of data needed to represent the operation of the hydraulic system without significant loss of information \cite{reduction_PHD}. The basic types of network components are maintained but the individual network components are combined and replaced. Skeletonization is not a single process but several different low-level element removal processes that must be applied in series in order to ensure that the demands in the network are logically reduced back to their source of supply. 

(in progress)

\emph{Comments for the supervisors:}

\textcolor{blue}{The plan is to try skeletonization in the first place. Write algorithm for serial/parallel pipe replacement and tree reduction. The Randers WSS should be split into different zones: (HNP, High zone, Low zone,  waterworks on the eastern side) and skeletonization should be carried out on each sub-system separately. After it is done, the sub-systems can be attached again. The HNP region and Low zone and the waterworks on the eastern side are mainly tree structures, they can be reduced significantly,i think, with skeletonization. I am not sure about the grid in the High zone. If the network after the skeletonization is not sufficiently reduced, then try a different method on the already skeletonized network.}

\todo{Q1: Verification of reduced networks}
\todo{Q2: Distributed optimization without reduction, large-scale network}

\subsection{Parameter fitting}
\label{parameter_fitting}

(in progress)

\subsection{Graph decomposition}
\label{graph_decomposition}

(in progress)

\subsection{Variable elimination}
\label{variable_elimination}

(in progress)

\section{Sectioning of the Randers WSS}
\label{sectioning_of_the_randers_WSS}

(in progress)

















